---
title: "final_project"
author: "Thomas W. Jones"
date: "4/18/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Define the model

$y \sim multinom(n,\theta) \implies \mathcal{L}(y|\theta,\alpha,\beta) \propto \prod_k \theta_k^{y_k}$

$\theta \sim Dir(\vec{\alpha}) \implies \pi(\theta) = \mathcal{B}(\vec{\alpha})\prod_k \theta_k^{\alpha_k - 1}$

$\alpha_k \sim Pareto(1, \beta) \implies \pi(\vec{\alpha}) = \prod_k\beta\alpha_k^{-(\beta + 1)}$

$\pi(\beta) \propto 1$

# Derive posterior distributions

\begin{align}
  P(\theta,\alpha,\beta|y) 
    &\propto \left[\prod_k \theta_k^{y_k}\right] 
      \left[\mathcal{B}(\vec{\alpha})\prod_k \theta_k^{\alpha_k - 1}\right]
      \left[\prod_k\beta\alpha_k^{-(\beta + 1)}\right]\\
    &= \beta^K\mathcal{B}(\vec{\alpha})
      \prod_k\theta_k^{y_k+\alpha_k - 1}\alpha_k^{-(\beta + 1)}\\
  P(\theta|\alpha,\beta,y)
    &\propto \prod_k \theta_k^{y_k+\alpha_k - 1}\\
    &\implies \theta|\alpha,\beta,y \sim Dir(\vec{y} + \vec{\alpha})\\
  P(\alpha|\theta,\beta,y) 
    &\propto \mathcal{B}(\vec{\alpha})
      \prod_k \theta_k^{y_k+\alpha_k - 1}\alpha_k^{-(\beta + 1)}\\
    &\implies \text{unknown distribution}\\
  P(\beta|\theta,\alpha,y)
    &\propto \beta^K(\prod_k\alpha_k)^{-(\beta + 1)}\\
    &\propto \beta^K(\prod_k\alpha_k)^{-\beta}\\
    &\implies \text{unknown distribution}
\end{align}


# Write a sampler

```{r}

library(textmineR)
library(gtools)

docs <- nih_sample$ABSTRACT_TEXT
names(docs) <- nih_sample$APPLICATION_ID

dtm <- CreateDtm(docs, stopword_vec = c(), verbose = FALSE)

y <- colSums(dtm)

B <- 100

theta <- matrix(0, nrow = B, ncol = length(y))

theta[ 1 , ] <- y / sum(y)

alpha <- matrix(0, nrow = B, ncol = length(y))

alpha[ 1, ] <- 0.01

acc_alpha <- numeric(B)

f_alpha <- function(x, beta, theta, y){
  # exp(out)
  
  # (gamma(sum(x)) / prod(gamma(x))) * prod(theta ^ (y + x - 1) * x ^ (-(beta - 1)))
  
  out <- (lgamma(sum(x)) - sum(lgamma(x))) + 
    sum((y + x - 1) * log(theta) + (-(beta - 1)) * log(x))
  
  out
}

beta <- numeric(B)

beta[ 1 ] <- 0.5

acc_beta <- numeric(B)

f_beta <- function(x, alpha) {
  # exp(out)
  out <- length(alpha) * log(x) - x * sum(log(alpha))
  
  out
}

for (j in 2:B) {
  
  # sample theta
  theta[ j , ] <- rdirichlet(1, y + alpha[ j - 1 , ] - 1)
  
  # sample alpha
  alpha_star <- rbeta(1, 2, 2)
  
  u <- runif(1)
  
  r <- (p(x = alpha_star, sigma2 = sigma2, mu = mu) / 
          p(x = alpha[ j - 1 ], sigma2 = sigma2, mu = mu) ) /
    (dbeta(alpha_star, 2, 2)/ dbeta(alpha[ j - 1 ], 2, 2)) 
    
  
  if (u < min(1, r)) {
    alpha[ j ] <- alpha_star
    acc[ j ] <- 1
  } else {
    alpha[ j ] <- alpha[ j - 1 ]
    acc[ j ] <- 0
  }

  
  # sample beta
  
  
}



```


